---
title: "Calculating NCI Comorbidity Index"
author: "Miriam Hu and Naichen Ni"
date: "Dec. 5, 2019"
output: ioslides_presentation
---

```{r setup, include = FALSE}
library(knitr)
  opts_chunk$set(echo = FALSE, cache = TRUE, autodep = TRUE, 
                 message = FALSE, warning = FALSE)
library(tidyverse)
library(plyr)
library(comorbidity)
library(readxl)
```

## Introduction

**Charlson Comorbidity Index:** most widely used measure of comorbid diseases

- First created in 1987
- Ranges from 0 to 24
- Each comorbid disease is assigned to a weight according to risk of death.
- e.g. Diabetes or chronic liver disease add 1 point to the score, while metastatic cancer or AIDS contribute 6 points.
- R package **comorbidity** can calculate the Charlson.

What if your dataset only had cancer patients?

**NCI Comorbidity Index:**

- Excludes tumors, leukemias, and lymphomas
- Slightly different weighting scheme (e.g. kidney disease has a score of 1.60 instead of 2)

## Example Data

```{r example-data}
# Clean the example data and write to csv for presentation -----
# example <- read_csv("example_data.csv")
# example <- example %>% subset(DIAGNOSIS_CODE_TYPE == "ICD-10")
# example <- example %>% subset(select = -DIAGNOSIS_CODE_TYPE)
# example$id <- sample(example$PATIENT_ID)
# example <- example %>% subset(select = -PATIENT_ID)
# example <- example %>% select(id, everything())
# example <- example %>% arrange(id)
# write_csv(example, "example_data.csv")

example <- read_csv("example-data.csv")
kable(example[1:3, ])
```

